### 2012-12-07 docs/paper.tex
Generalized model to allow for more probability distributions

\documentclass[a4paper,12pt]{article}
\usepackage{amsmath} % For 'cases' curly brace
\usepackage{amsfonts} % For bold number set font
\begin{document}
\title{Mixed Non-Bayesian and DeGroot Learning}
\author{Pedro d'Aquino\\pdaquino@umich.edu \and Dan Clark\\ddclark@umich.edu}
\date{December 11, 2012}

\maketitle

\section{Introduction}

\cite{Jadbabaie2012}
\cite{DeGroot1974}

\section{The Model}

\subsection{The Network}

(Mixed Non-Bayesian and DeGroot nodes)

\subsection{States and Signals}

We will consider the state of the world to be a probability distribution $p^*(t) \in \mathbb{P}$.  $\mathbb{P}$ is a pre-defined class of distributions where each  $p^*(t) \in \mathbb{P}$ is distinguishable by a single parameter.  This single parameter $m^*$ where $\{m^* \in \mathbb{R}: m_{min} <= m^* <= m_{max} \}$  will be the state which the nodes in the network are attempting to learn.  That is, $m^*$ is the ``true'' state of the world.  \emph{A priori}, the nodes know the class of distribution but not the specific value of $m^*$; only that it lies uniformly in the range $[m_{min}, m_{max}]$.

Our model allows for $\mathbb{P}$ to be any class of probability distribution as long as it can be completely specified by a single parameter $m^*$.  Examples include the Poisson distribution with parameter $m^*$ and Gaussian distribution with mean $m^*$ and constant variance.  We also introduce a trapezoidal probability distribution in section~\ref{sec:trapezoidal_distribution} which serves as a valid choice for $\mathbb{P}$.

At each timestep $t$, every Non-Bayesian node $i$ will receive a signal $s_{i,t}$ which is a value drawn from $p^*(x)$.  These observations will be used by the nodes in the network to construct a belief about $m^*$.

\subsubsection{The Trapezoidal Distribution}
\label{sec:trapezoidal_distribution}

As an example of a class of distributions which can be used in our model we present the following trapezoidal distribution.

\begin{equation}
p^*(x)=\begin{cases}
m^* \cdot x + \frac{1}{2}(2-m^*) & 0 \le x \le 1 \\
0, & \text{otherwise}
\end{cases}
\end{equation}

% @todo[DDC] Graph/figure of this would probably be good to have.

The slope $\{ m^* \in \mathbb{R} : m_{min} = -2 \le m^* \le 2 = m_{max} \}$ will be the defining characteristic of the distribution.  We have chosen to bound $m^*$ between -2 and 2 and add the term $\frac{1}{2}(2-m^*)$ so that the support of $p^*(x)$  is always over $[0,1]$.

\subsection{Representing Belief States}

For a DeGroot node $i$, we represent the belief state of $i$ at time $t$ as a single value 
$m_{i,t} \in \mathbb{R}$.  However, belief of Non-Bayesian nodes is a probability distribution over a finite set of possible states of the world.  There is thus a tension between these two schemes and it is necessary to develop a conversion between the belief states of Non-Bayesian and DeGroot nodes.

\subsubsection{Belief in Non-Bayesian Nodes}

To represent a belief about the continuous slope value $m^*$ for a Non-Bayesian node we discretize $m$ in the following manner.  Let $\Theta = \{\theta_0, \theta_1,...\theta_n\}, n < \infty$ be the discrete set of possible states of the world over which each Non-Bayesian node holds a distribution of belief. Each $\theta_k$ corresponds to the belief that the slope of $p^*(x)$ is equal to $m_k$ where
\begin{equation}
\label{eq:theta_meaning}
\hat{m_k} = \frac{(m_{max} - m_{min})k}{n - 1} + m_{min}
\end{equation}
As an example, if we take $n = 21$ belief states, $m_{min} = -2$, and $m_{max} = 2$, then we have
\begin{equation}
\nonumber
\begin{cases}
& \theta_0: m^* = -2.0 \\
& \theta_1: m^* = -1.9 \\
& ... \\
& \theta_{20}: m^* = 2.0
\end{cases}
\end{equation}
For each Non-Bayesian node $i$, the belief at time $t$ that $\theta_k$ is the true state of the world is denoted by $\mu_{i,t}(\theta_k)$.  Thus $\{ \mu_{i,t}(\theta_1), \mu_{i,t}(\theta_2), ... \mu_{i,t}(\theta_n) \}$ is a probability distribution over the set of world states for fixed $i,t$.

\subsection{Learning in DeGroot Nodes}
DeGroot nodes in our model do not directly receive signals; they act merely as message-passers in the network.  For the most part our DeGroot nodes behave in a fashion similarly to nodes in a standard DeGroot-style network.  That is, a DeGroot node's $i$'s belief of the true slope value $m^*$ is given at time $t+1$ as
\begin{equation}
\label{eq:degroot_update}
m_{i,t+1} = \sum_{j \in N(i)} a_{ij}m_{j,t}
\end{equation}
where $N(i)$ denotes the neighbors of $i$.  $a_{ij}$ represents the level of ``trust'' that a DeGroot node $i$ has in neighbor $j$, such that

\begin{equation}
\nonumber
\sum_{j \in N(i)} a_{ij} = 1
\end{equation}
for all DeGroot nodes $i$.  Thus the DeGroot update in equation \ref{eq:degroot_update} is simply a weighted average of its neighbors' beliefs in the previous timestep.

When a DeGroot node with a Non-Bayesian neighbor performs an update it must obtain a value 
$m_{j,t}^\prime$ for the belief of this neighbor.  This is calculated as

\begin{equation}
m_{j,t}^\prime = \sum_{k=1}^n\mu_{j,t}(\theta_k)\hat{m_k}
\end{equation}

Here we are taking a weighted average of the $\hat{m_k}$ slope values represented by each state $\theta_k$ as defined in equation~\ref{eq:theta_meaning}, where the weights are given by $j$'s level of belief in each state.

\subsection{Learning in Non-Bayesian Nodes}

The update of a Non-Bayesian node in our model is handled similarly to the method laid out 
in \cite{Jadbabaie2012}.  We denote the belief of a node $i$ that it will receive the signal $s_i$ at time $t$ as $m_{i,t}(s_i)$, defined as follows:

\begin{equation}
m_{i,t}(s_i) = \int_\Theta \ell_i(s_i|\theta)d\mu_{i,t}(\theta) = \sum_{k=1}^n \ell_i(s_i|\theta_k)\mu_{i,t}(\theta_k)
\end{equation}

% @todo[DDC] More detail for derivation/explanation of likelihood?
The likelihood function $\ell_i(s_i|\theta_k)$ can be obtained from the probability distribution function represented by the belief state $\theta_k$.  As an example, for the trapezoidal distribution described in section~\ref{sec:trapezoidal_distribution}, the likelihood function would be

\begin{equation}
\label{eq:likelihood}
\ell_i(s_i|\theta_k) \propto \hat{m_k} s_i + \frac{1}{2}(2 - \hat{m_k})
\end{equation}
where $\hat{m_k}$ is defined as in equation~\ref{eq:theta_meaning}.

The update of a node's belief in each state $\theta_k$ for a given time period $t$ is then given by

\begin{equation}
\label{eq:non_bayesian_update}
\mu_{i,t+1}(\theta_k) = a_{ii}\mu_{i,t}(\theta_k)\frac{\ell_i(\omega_{i,t+1}|\theta_k)}{m_{i,t}(\omega_{i,t+1})} + \sum_{j \in N(i)} a_{ij}\mu_{j,t}(\theta_k)
\end{equation}

Here the first term is the Bayesian update of the belief $\mu_{i,t}(\theta_k)$ after observing the signal $\omega_{i,t+1}$, multiplied by the node's self reliance $a_{ii}$.  The summation is the linear incorporation of the beliefs of $i$'s neighbors.

This update differs from the previous work in two major ways.  Firstly, we provide each Non-Bayesian node with only a single signal per timestep, where this signal is a single draw from the probability distribution $p^*(x)$.

Secondly, we must define $\mu_{j,t}(\theta_k)$ when $j$ is a DeGroot node.  We consider two methods for doing so, defined in sections \ref{sec:draw_from_degroot} and \ref{sec:likelihood_metric} respectively.

\subsubsection{Belief Distribution from Draw of Probability Distribution}
\label{sec:draw_from_degroot}

We require a method for converting a DeGroot node $j$'s belief $m_{j,t} \in \mathbb{R}$ to a series of priors which can be included in a Non-Bayesian node $i$'s belief update.  The first method we consider involves generating a probability distribution $p_{j,t}$ from $j$'s belief at time $t$ and drawing from this distribution.  The generated distribution is the $p_{j,t} \in \mathbb{P}$ where $j$'s belief $m_{j,t}$ is the single parameter defining the distribution.  If $\mathbb{P}$ is the trapezoidal distribution defined in section~\ref{sec:trapezoidal_distribution}, then for each DeGroot neighbor of $i$ at time $t$ we have

\begin{equation}
p_{j,t}(x)=\begin{cases}
m_{j,t} x + \frac{1}{2}(2-m_{j,t}) & 0 \le x \le 1 \\
0, & \text{otherwise}
\end{cases}
\end{equation}

This is simply the distribution which $j$ believes to be the state of the world.  We then draw a single signal $s_{j,t}^\prime$ from this distribution.  We create a belief distribution $\mu_{j,t}^\prime$ from $s_{j,t}^\prime$ by performing a Bayesian-style update with equal priors.

\begin{equation}
\mu_{j,t}^\prime(\theta_k) = \frac{\ell_i(s_{j,t}^\prime|\theta_k)}{\sum_{i=1}^{n} \ell_i(s_{j,t}^\prime|\theta_i)}
\end{equation}

This generated belief distribution $\mu_{j,t}^\prime$ is then used for the values of $\mu_{j,t}(\theta_k)$ for the DeGroot node $j$ in the Non-Bayesian update of equation~\ref{eq:non_bayesian_update}.

\subsubsection{Belief Distribution from Likelihood Distance Metric}
\label{sec:likelihood_metric}

The second method we consider for creating a probability distribution over the set of belief states $\Theta$ is to define an exponential distance metric such that

\begin{equation}
\mu_{j,t}^\prime(\theta_k) \propto e^{-(m_{j,t}-\hat{m_k})^2}
\end{equation}
where $\mu_{j,t}^\prime$ is the belief distribution which is used for the DeGroot node $j$'s values of $\mu_{j,t}(\theta_k)$ in the Non-Bayesian update of equation~\ref{eq:non_bayesian_update}.

Since $\mu_{j,t}^\prime$ must be a probability distribution we must normalize the values for each state with the value

\begin{equation}
q_{j,t} = \sum_{k=1}^n e^{-(m_{j,t}-\hat{m_k})^2}
\end{equation}

Thus the calculation to create the DeGroot node's level of belief in each state $\theta_k$ is given by

\begin{equation}
\mu_{j,t}^\prime(\theta_k) = e^{-(m_{j,t}-\hat{m_i})^2} / q_{j,t}
\end{equation}

\bibliographystyle{plain}
\bibliography{eecs598Report}

\end{document}